---
title: "Visualizing the History of Nobel Winners"
author: "Ceren Ünal"
date: "2024-07-13"
output: 
  html_document:
    #code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
    toc_collapsed: false
    theme: flatly
    number_sections: yes
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import Data

```{r}
library(tidyverse)
```

```{r}
nobel <- read_csv("nobel.csv")
```

# Explore Data

```{r}
glimpse(nobel)
```

The data set consists of 18 variables and 1000 observations. Many are character variables, with only 2 being numeric and 2 being dates. 

```{r}

min(nobel$year)
max(nobel$year)
```

The data covers years from 1901 to 2023.  

```{r}

colSums(is.na(nobel))
```

10 variables have missing values, including sex variable. 

```{r}
nobel %>% 
  filter(is.na(sex)) %>% 
  mutate(birth_date_na = is.na (birth_date)) %>% 
  summarise(sum(birth_date_na))
```

All observations that have a missing value in the sex variable, are also missing the birth_date values. It seems that these are all organizations, not persons.  

```{r}
length(unique(nobel$category))
length(unique(nobel$full_name))
length(unique(nobel$organization_name))

```

There are 6 prize categories.  
There have been 993 unique winners, meaning some have won more than once.  
There are also only 326 organization names, some have received multiple prizes.  



# What is the most commonly awarded gender and birth country?  


```{r}
nobel %>%
  group_by(sex) %>%
  count() %>%
  arrange(desc(n))
```

The most commonly awarded gender is male. 


```{r}
nobel %>%
  group_by(birth_country) %>%
  count() %>%
  arrange(desc(n))
```

The most commonly awarded country is the USA.  

# What decade had the highest proportion of US-born winners?  


```{r}
nobel %>% 
  mutate(decade = year %/% 10 * 10) %>% 
  group_by(decade) %>% 
  mutate(all_winners = n()) %>% 
  group_by(decade, birth_country) %>% 
  filter(birth_country == "United States of America") %>% 
  mutate(us_born_winners = n()) %>% 
  ungroup() %>% 
  group_by(decade) %>% 
  reframe(prop = us_born_winners/all_winners) %>% 
  arrange(desc(prop))
```

2000s had the highest proportion of US-born winners. 

# What decade and category pair had the highest proportion of female laureates?  

```{r}
nobel %>% 
  select(year, category, sex) %>% 
  mutate(decade = year %/% 10 * 10) %>% 
  group_by(decade, category) %>% 
  mutate(all_winners = n()) %>% 
  group_by(decade, sex) %>% 
  filter(sex == "Female") %>% 
  mutate(female_winners = n()) %>% 
  ungroup() %>% 
  group_by(decade, category) %>% 
  reframe(prop = female_winners/all_winners) %>% 
  arrange(desc(prop))
```

2020s and Literature pair had the highest proportion of female winners. 


# Who was the first woman to receive a Nobel Prize, and in what category?  

```{r}
nobel %>% 
  select(year, sex, category) %>% 
  filter(sex == "Female") %>% 
  arrange(year)
```

The first woman to receive the Nobel Prize, was awarded the prize in 1903 in physics category.  

# Which individuals or organizations have won multiple Nobel Prizes throughout the years? 

```{r}
nobel %>% 
  group_by(full_name) %>% 
  count() %>% 
  filter(n > 1) %>% 
  arrange(desc(n))
```

2 organizations and 4 individuals were awarded multiple Nobel Prizes.  
